---
title: "p8131_hw5_mp3745"
author: "Matthew Parker"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(nnet)
library(MASS)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

#### (a)

Read in the crab data
```{r}
crab_data = read.table("./data/HW5-crab.txt", header = TRUE)
```


Fit a Poisson model (m1) with log link with W as the single predictor
```{r}
m1 = glm(Sa ~ W, 
         family = poisson, 
         data = crab_data) 

# View summary of m1
summary(m1)

# Check goodness of fit of M1
m1_dev = sum(residuals(m1, type = 'deviance')^2)

m1_pval = 1 - pchisq(m1_dev, (nrow(crab_data) - 2))
```

Since p- value = `r m1_pval` < 0.05, we reject the null hypothesis of the model being a good fit for the data. The log ratio of the number of satellites for a 1 unit increase in carapace width is `r m1$coefficients[2]`. The log of the expected number of satellites for a crab with a zero carapace width is `r m1$coefficients[1]`.


#### (b)

Fit a model (m2) with W and Wt as predictors, then compare to m1
```{r}
m2 = glm(Sa ~ W + Wt, 
         family = poisson, 
         data = crab_data) 

# View summary of m2
summary(m2)

# Compare m2 to m1
m1_m2_dev_stat = m1$deviance - m2$deviance
m1_m2_df = (nrow(crab_data) - 2) - (nrow(crab_data) - 3) 
m1_m2_pval = 1 - pchisq(m1_m2_dev_stat, df = m1_m2_df) 
```

Since the p-value = `r m1_m2_pval` < 0.05, we reject the null hypothesis that the smaller model fits the data as well as the larger model. This means the smaller model has a lack of fit and the larger model is better to use. In m2, the log ratio of the number of satellites for a 1 unit increase in carapace width is `r m2$coefficients[2]`, holding weight constant. The log ratio of the number of satellites for a 1 unit increase in weight is `r m2$coefficients[3]`, holding carapace width constant. The log of the expected number of satellites for a crab with a zero carapace width and zero weight is `r m2$coefficients[1]`.


#### (c)

Check over dispersion in M2
```{r}
# Residuals
m2_res = residuals(m2, type = 'pearson')

m2_resid_points = tibble(
  x = qnorm((173 + 1:173 + 0.5)/(2 * 173 + 1.125)),
  y = sort(abs(m2_res))
)

m2_resid_points %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point() + 
  geom_abline(slope = 1) + 
  scale_x_continuous(limits = c(0, 8)) +
  scale_y_continuous(limits = c(0, 8)) +
  labs(
    x = "Expected Half-Normal Order Stats",
    y = "Ordered Abs Pearson Residuals"
  )
```

Based on the above plot, there may be overdispersion. There is clear deviance from the reference line.

m2 with overdispersion
```{r}
# Estimate overdispersion parameter
g = sum(m2_res^2)

phi = g / (173 - 3)

# m2 with overdispersion parameter
summary(m2, dispersion = phi)
```

In the updated model, our estimates and their interpretations are still valid. In the updated model, the std. error has been inflated by `r phi`. This has caused the z value and Pr(>|z|) to change. The coefficients for the both carapace width and weight in the model have now clearly become insignificant, indicating that:

* there may not be a significant association between carapace width with number of satellites, when accounting for weight, and

* there may not be a significant association between weight with number of satellites, when accounting for carapce width.


## Problem 2

#### (a)

Read in the parasite data
```{r}
par_data = read.table("./data/HW5-parasite.txt", header = TRUE) %>% 
  janitor::clean_names() %>% 
  mutate(
    area = as_factor(area),
    year = as_factor(year)
  )
```


Fit a Poisson model with log link to the data with area, year, and length as predictors
```{r}
par_fit = glm(intensity ~ area + year + length, 
         family = poisson, 
         data = par_data) 

# View summary of par_fit
summary(par_fit)
```

The log ratio of the number of parasites for a fish with area equal to 2 vs area equal to 1 is `r par_fit$coefficients[2]`, holding year and length constant. The log ratio of the number of parasites for a fish with area equal to 3 vs area equal to 1 is `r par_fit$coefficients[3]`, holding year and length constant. The log ratio of the number of parasites for a fish with area equal to 4 vs area equal to 1 is `r par_fit$coefficients[4]`, holding year and length constant. The log ratio of the number of parasites for a fish of year 2000 vs year 1999 is `r par_fit$coefficients[5]`, holding area and length constant. The log ratio of the number of parasites for a fish of year 2001 vs year 1999 is `r par_fit$coefficients[6]`, holding area and length constant. The log ratio of the number of parasites for a 1 unit increase in length is `r par_fit$coefficients[7]`, holding area and year constant. The log of the expected number of parasites for a fish with area equal to 1 and of year 1999 with length equal to zero is `r par_fit$coefficients[1]`.


#### (b)

Test for goodness of fit
```{r}
# Check goodness of fit of par_fit
par_fit_dev = sum(residuals(par_fit, type = 'deviance')^2)

par_fit_pval = 1 - pchisq(par_fit_dev, (nrow(par_data) - 7))
```

Since p-value = `r par_fit_pval` < 0.05, we reject the null hypothesis of the model being a good fit for the data.